---
title: "Analysis of the World dataset"
output:
  html_document: default
  pdf_document: default
---

## 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install the packages
```{r eval=FALSE}
# r setup, eval=FALSE
install.packages('tidyverse')
install.packages('plotrix')
install.packages('ggplot2')
install.packages("dplyr")

```


## Load the Packages
```{r}
library(tidyverse)  # package for data analysis in R
library(plotrix)  # contains various functions for plotting
```



## 1. Load Data

Load the world.csv dataset and store it as an object "world.data".

```{r}
world.data <- read_csv('world csv new.csv')
# view first few rows
head(world.data)
```



## 2. Create frequency table of the OECD variable

```{r}
# create frequency table with the table() function
freq_table <- table(world.data$oecd)
# calculate the percentages
percentage <- freq_table / sum(freq_table) * 100
# convert the frequency table to a dataframe
ft.oecd <- as.data.frame(freq_table)
# add the Percentage column to the dataframe
ft.oecd$Percentage <- percentage
# rename the first column to "OECD Member?"
names(ft.oecd)[1] <- "OECD Member?"
ft.oecd
```
 
 
 
## 3. Frequency Table Observation

* 30 countries are OECD members.
* 161 countries are not OECD members.
* 15.7% of the countries are OECD members.
* 84.3% of the countries are not OECD members.



## 4. Barchart of the OECD variable
Use the ggplot library in tidyverse to create a bar chart

```{r}
ggplot(data=world.data) +
  geom_bar(mapping=aes(x=oecd), fill='darkblue') + 
  # Add title to the plot
  labs(title='How Many Countries Have OECD Membership?') + 
  # centralize the title
  theme(plot.title = ggplot2::element_text(hjust = 0.5)) + 
  # Add xlabel to the x-axis
  xlab('OECD membership') +
  # Add ylabel to the y-axis
  ylab('Number of countries')
```



## 5. OECD member states and non-democratic states
#### OECD member states
- Australia
- Austria
- Belgium

#### Non-democratic states
- Afghanistan
- Algeria
- Angola



## 6. Statistical Description of the `gdp_10_thou` variable.

```{r}
# print summary statistic of the `gdp_10_thou` variable
summary(world.data$gdp_10_thou)
```
  
  
```{r}
# check the standard deviation & ignore missing values
sd(world.data$gdp_10_thou, na.rm=TRUE)
```

* The minimum value is 0.0090
* The maximum value is 4.7354
* The median value is 0.1897
* The mean value is 0.6018
* The 1st quartile value is 0.0503
* The 3rd quartile value is 0.6320
* The standard deviation is 0.9433982



## 7. Left or Right Skewed?

Since the mean is greater than the median ($6,018 > $ 1,897), the data is positively skewed (skewed to the right).



## 8. Histogram  of the per capita GDP variable

```{r warning=FALSE}
# r setup, warning=FALSE
ggplot(data=world.data) +
  # Plot histogram
  geom_histogram(mapping=ggplot2::aes(x=gdp_10_thou), fill='darkblue', binwidth=0.4) + 
  labs(title='Histogram of the Per Captia GDP Variable') + 
  # centralize the title
  theme(plot.title=ggplot2::element_text(hjust = 0.5)) + 
  # Add xlabel to the x-axis
  xlab('Per capita GDP (in 10,000 US dollars)') +
  # Add ylabel to the y-axis
  ylab('Number of countries')
```



## 9. The two countries with per capita GDP over $40,000

```{r}
# filter dataframe to view countries whose GDP is over $40,000
world.data %>% filter(gdp_10_thou>4)
```

* The two countries with per capita GDP over $40,000 are Luxembourg and Norway.



## 10.  Calculate the standard error of the per capita GDP variable.

```{r warning=FALSE}
# r setup, warning=FALSE
# use the plotrix package to calculate the standard error
print(std.error(world.data$gdp_10_thou))

# or manually calculate it with the formular: standard deviation/square root of the sample size
# sample size is the total number of rows minus the number of missing values in the gdp_10thou variable

standard_error <- sd(world.data$gdp_10_thou, na.rm=TRUE)/sqrt(nrow(world.data)-sum(is.na(world.data$gdp_10_thou)))
print(standard_error)
```

The standard error is 0.07091015



## 11. Calculate the 95% Confidence Interval

Formula: 
Confidence Interval = sample mean +- margin of error
margin of error = z* x standard error

For 95% confidence interval, z* = 1.96

Therefore, 95% confidence interval = sample mean +- 1.96 * stardard error 


```{r}
# calculate the margin of error using the standard error
margin_of_error <- 1.96*standard_error
# calculate the mean of the gdp_10_thou variable
sample_mean <- mean(world.data$gdp_10_thou, na.rm=TRUE)

# Print out the results
cat("The margin of error is", margin_of_error, '\n')
cat("The sample mean is", sample_mean, '\n')

# Calculate the confidence interval range
confidence_interval_upper <- sample_mean + margin_of_error
confidence_interval_lower <- sample_mean - margin_of_error

# Print out the results
cat("The upper confidence interval is", confidence_interval_upper, '\n')
cat("The lower confidence interval is", confidence_interval_lower, '\n\n')
cat("The 95% confidence interval ranges from", confidence_interval_lower, '-', confidence_interval_upper)
```



## 12. Histograms of per capita GDP variable. One for democracies and the other for non-democracies.

```{r warning=FALSE}
# r setup, warning=FALSE
ggplot2::ggplot(data=world.data) +
  ggplot2::geom_histogram(mapping=ggplot2::aes(x=gdp_10_thou), fill='darkblue', binwidth=0.4) + 
  ggplot2::labs(title='Histogram of the Per Captia GDP Variable For Democracies and Non-democracies ') + 
  # centralize the title
  ggplot2::theme(plot.title = element_text(hjust = 0.5)) + 
  ggplot2::xlab('Per capita GDP (in 10,000 US dollars)') +
  ggplot2::ylab('Number of countries') + 
  # create a facet_wrap of the democ_regime variable
  ggplot2::facet_wrap(~democ_regime)
```



## 13. Fix Histogram

I have to get rid of the missing values in the democ_regime variable represented by 'NA'. Then create a new
variable called 'dem.dum' based on the variable 'democ_regime'. It will be more descriptive than the 'democ_regime' variable.


```{r}
# remove NA values in the democ_regime variable
dem.gdp <- world.data %>% drop_na('democ_regime')
# create a new column called dem.dum in the dem.gdp dataframe
# the value will be 'Democracy' if the value in the democ_regime variable is 'Yes'
# otherwise it will be 'Autocracy'
dem.gdp <- transform(dem.gdp, dem.dum=ifelse(democ_regime=='Yes', 'Democracy', 'Autocracy'))
# view a few rows of the new dataframe
head(dem.gdp)
```

Now plot the histogram with this cleaned dataframe.

```{r warning=FALSE}
# r setup, warning=FALSE
ggplot(data=dem.gdp) +
  geom_histogram(mapping=aes(x=gdp_10_thou), fill='darkblue', binwidth=0.4) + 
  labs(title='Histogram of the Per Captia GDP Variable For Democracies and Non-democracies ') + 
  # centralize the title
  theme(plot.title = element_text(hjust=0.5)) + 
  xlab('Per capita GDP (in 10,000 US dollars)') +
  ylab('Number of countries') + 
  facet_wrap(~dem.dum)
```



## 14. Mean of per capita GDP for democracies & the 95 % confidence interval

```{r}
# filter data to get democratic countries only
democracies <- dem.gdp %>% filter(dem.dum=='Democracy')
head(democracies)
```


```{r}
# create a function to calculate the mean and the 95% confidence interval
# the function will take in a dataframe as an argument
mean_and_95_CI <- function(dataframe) {
  # calculate the standard error
  standard_error <- std.error(dataframe$gdp_10_thou)
  # calculate the margin of error
  margin_of_error <- 1.96 * standard_error
  # calculate the sample mean
  sample_mean <- mean(dataframe$gdp_10_thou, na.rm=TRUE)
  
  # Display results
  cat("The standard_error is", standard_error, '\n')
  cat("The margin of error is", margin_of_error, '\n')
  cat("The sample mean is", sample_mean, '\n')
  
  # Calculate the confidence interval range
  confidence_interval_upper <- sample_mean + margin_of_error
  confidence_interval_lower <- sample_mean - margin_of_error
  
  # Print out the result
  cat("The upper confidence interval is", confidence_interval_upper, '\n')
  cat("The lower confidence interval is", confidence_interval_lower, '\n\n')
  cat("The 95% confidence interval ranges from", confidence_interval_lower, '-', confidence_interval_upper)
}

# invoke function for democratic countries 
# by passing the democracies dataframe
mean_and_95_CI(democracies)

```



## 15.  Mean of per capita GDP & the 95 % Confidence Interval For Autocracies

```{r}
# filter data to get autocratic countries only
autocracies <- dem.gdp %>% filter(dem.dum=='Autocracy')
head(autocracies)
```

```{r}
# invoke function for autocratic countries
# by passing the autocracies dataframe
mean_and_95_CI(autocracies)
```


* The mean of per capita GDP for democracies is 0.8013927.
* The 95% confidence interval for democracies ranges from 0.5984288 - 1.004357.


* The mean of per capita GDP for autocracies (non-democracies) is 0.2819132.
* The 95% confidence interval for autocracies ranges from 0.1549886 - 0.4088379.

